---
title: "P. taeda_snesitivity_and_uncertainty_analysis"
author: "David_Herrera"
date: "10/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Markdown present the code used for calculating the parameters related to the model presented in the Figure 2 of the manucript "Probability distributions of non-structural carbon ages and transit times provide insights in carbon allocation dynamics of mature trees" for the species P. taeda. The parameters are calculated implementing the model "Allometrically Constrained Growth and Carbon Allocation" ACGCA (Ogle and Pacala 2009). Here it is also presented the code for running the analysis of the uncertainties related to the model parameters in the model output (mean ages and mean transit times) and the sensitivity of the mean ages and transit time to numerical changes in those parametes. 

### packages needed 
```{r}
library(SoilR)
library(boot)
library(FME) 
library(sensitivity)
```

set the directory where all the auxiliary files are stored. The functions for running the ACGCA model are in the sicript "script_functions_model_ogleV2.R" 

```{r}
setwd("/Users/_dherrera/NSC_ages_and_transit_times/NSC_ages_and_transit_times/code/P_taeda/")
source("/Users/_dherrera/NSC_ages_and_transit_times/NSC_ages_and_transit_times/code/script_functions_model_ACGCA.R")

```

## P. taeda parameter calculation 

The parameters provided for Ogle and Pacala (2009) are written in the data frame paramogle. With this parameters we ran the model ACGCA and estimated the parameter for the model presented in the figure 2 of the manuscript. the data frame was contructed with the mean values for P. taeda (first row) and A. rubrum (second row). From here on we will just use the parameters corresponding to P. taeda. 


```{r}

paramogle=data.frame(Crwparam=c(0.308, 0.365),
                     Radeff=c(4.5, 6.75),
                     Pmax=c(2100, 2100),
                     Cgl=c(1.51, 1.45),
                     Cgr=c(1.30, 1.25),
                     Crw=c(1.47, 1.37),
                     Rml=c(0.95, 1.25),
                     Rmr=c(0.75, 0.75),
                     Rms=c(0.025, 0.025),
                     Sl=c(0.33, 1),
                     So=c(0.05, 0.05),
                     Sr=c(0.5, 1),
                     dl=c(0.11, 0.095),
                     dr=c(0.08, 0.15),
                     Stmaxspw=c(265000, 131000),
                     denspw=c(6.67e-7, 6.67e-7),
                     Xylspwrate=c(0.62, 0.12),
                     k=c(0.55, 0.7),
                     Maxcrwdepth=c(0.95, 0.95),
                     pa=c(380000, 525000),
                     pb=c(380000, 525000),
                     pr=c(200000, 160000),
                     rootrad=c(0.00027, 0.00015),
                     SLA=c(0.0060, 0.0141),
                     n=c(0.71, 0.64),
                     nb=c(0.045, 0.045),
                     fiH=c(220, 263),
                     lamb=c(0.95, 0.95),
                     Hmax=c(42, 27.5),
                     SWmax=c(0.06, 0.10),
                     f1=c(4, 4),
                     f2=c(2100, 7000), 
                     lambs=c(0.95, 0.95),
                     lambh=c(0.95, 0.95),
                     R0=c(1.434, 1.909),
                     R40=c(3.873, 5.592),
                     m=c(0.95, 0.95))

paraPtadea=paramogle[1,] ### here we choosed the parameters for P. taeda. 
```


Then we ran the ACGCA model for 800 years (niter) to assure that we selected the parameters at the steady state of the system. This simulation 
```{r}


 # r1=0.0001### inital value for radious
 # respPtadea=ACGCAtargetallom(r1=r1, dr=0, t=0, paraPtadea=paraPtadea, Cs=c(0))## initial conditions 
 # respPtadea4=modhealty(initcond = respPtadea, niter=800, param=paraPtadea, tao=0.000001, errori=10000)
 # respPtadea4=respPtadea4[-c(nrow(respPtadea4)),] ## the last row is incomplete then we eliminated.


  load("respPtaeda_PmaxV6.Rda")# The results of the previous simulation were saved in this file and can be laded here. 
 
```

Then we calculated the fluxes, the pool sizes and the transfer coefficients for the model in Figure 2

```{r}

## We took then the fluxes and stocks in stady state: the values at the point of 700 years of the simulations of the ACGCA model

flux300yr=respPtadea4[which(respPtadea4$t==700),] 


## With these values we proceeded to calculate the transfer coefficients for the model used in the manuscript and described in Figure 2

E=flux300yr$E+flux300yr$Rm ##g of glucose 

## fluxes for leave compartment 

PFlT=flux300yr$Fl*flux300yr$E; ## PFlT is the amount of g glucose allocated to leaves 
SsL=PFlT # SsL is the g of glucose in the fast cycling pool of sugars in the leaves
Gl=PFlT*(paraPtadea["Cgl"]/(paraPtadea["Cgl"]+paraPtadea["dl"])); Gl### grams of glucose per year invested in constructing leaves (flux of glucose to grow leaves)
FB=flux300yr$Bl## biomass (g dw) in the leaves 
biomlgdw=PFlT/(paraPtadea["Cgl"]+paraPtadea["dl"]); biomlgdw## grams of wood constructed from the allocated carbon to leaves
stl=PFlT*(paraPtadea["dl"]/(paraPtadea["Cgl"]+paraPtadea["dl"])); stl## grams of glucose per year stored in the leaves 
senlTgdw=paraPtadea["Sl"]*flux300yr$Bl; senlTgdw## g of wood lost in litter fall per year 
senlT=paraPtadea["Sl"]*flux300yr$Bl*paraPtadea["Cgl"] ## g of glucose per year lost in the litter fall 
dlSl=paraPtadea["Sl"]*flux300yr$Bl*paraPtadea["dl"]; dlSl ##g of glucose per year translocated to the E pool from the storage
FSNSC=flux300yr$Cl ## size of the storage pool in the leaves 


## fluxes for fine root compartment 
PFrT=flux300yr$Fr*flux300yr$E; PFrT## amount of g glucose allocated to roots per year 
Ssr=PFrT ## g glucose in the fast cycling pool of sugar in the roots
Gr=PFrT*(paraPtadea["Cgr"]/(paraPtadea["Cgr"]+paraPtadea["dr"])); Gr## g of glucose alocate to biomass in  roots
biomrgdw=PFrT/(paraPtadea["Cgr"]+paraPtadea["dr"]); biomrgdw## grams of wood constructed from the allocated carbon to roots
strr=PFrT*(paraPtadea["dr"]/(paraPtadea["Cgr"]+paraPtadea["dr"])); strr## grams of glucose per year stored in the roots 
senrTgdw=paraPtadea["Sr"]*flux300yr$Br## g of wood lost in senescence of fine roots 
senrT=paraPtadea["Sr"]*flux300yr$Br*paraPtadea["Cgr"] ## g of glucose lost in senescence of fine roots 
drSr=paraPtadea["Sr"]*flux300yr$Br*paraPtadea["dr"];drSr ## g of glucose reallocate to the transient comparment (E) from roots 
RSNSC=flux300yr$Cr## size of the sotrage compartment in roots


## fluxes for woody compartments 

PFtT=flux300yr$Ft*flux300yr$E; PFtT## amount of g glucose allocated to the main trunk
SsT=PFtT## g of glucose allocated to the fast cycling pool in the trunk 
Gt=PFtT*(paraPtadea["Crw"]/(paraPtadea["Crw"]+flux300yr$ds))#g of glucose per year allocated to grow wood in the trunk 
SB=flux300yr$Bts+flux300yr$Bth ## g of wood in the trunk (biomase of the main trunk)
strt=PFtT*(flux300yr$ds/(paraPtadea["Crw"]+flux300yr$ds))## grams of glucose per year allocated to storage in the trunk 

## fluxes for other wood compartment (branches and coarse roots)
PFoT=flux300yr$Fo*flux300yr$E; PFoT## amount of g glucose allocated to branches and coarse roots per year 
SsO=PFoT ## size of the fast cycling pool of branches and coarse roots  in g of glucose
Go=PFoT*(paraPtadea["Crw"]/(paraPtadea["Crw"]+flux300yr$ds))### grams of glucose per year allocated to the biomass of branches and coarse roots
Bo=flux300yr$Bos+flux300yr$Boh ## g of wood in the branches and coarse root
senoTgdw=PFoT/(paraPtadea["Crw"]+flux300yr$ds)#g of wood lost in literfall
senoT=PFoT/(paraPtadea["Crw"]+flux300yr$ds)*paraPtadea["Crw"]## g of glucose lost in literfall
stro=PFoT*(flux300yr$ds/(paraPtadea["Crw"]+flux300yr$ds))## g of glucose per year allocated to storage inthe wood
So=(PFoT/(paraPtadea["Crw"]+flux300yr$ds))/Bo## g of wood per year lost in litter fall from branches and coarse roots 
dsSo=flux300yr$ds*So*flux300yr$Bos### g of glucose per year retranslocated to E from branches and coarse roots 
dwSo=strt+stro-dsSo## g of glucose per year lost in litterfall from the carbon storage comparment 
SSNSC=flux300yr$Cs## glocose stored in wood of branches and coarse roots

### calculation of carbon transfer coefficients. These are the parameters of the model decribed in Fig 2, for the P. taeda case.

ratesPtadea=c(Rm=as.numeric(flux300yr$Rm/E),
              Fl=as.numeric(PFlT/E),
              Fr=as.numeric(PFrT/E),
              Ft=as.numeric(PFtT/E),
              Fo=as.numeric(PFoT/E),
              PGl=as.numeric(Gl/SsL),
              PGr=as.numeric(Gr/Ssr),
              PGo=as.numeric(Go/SsO),
              PGt=as.numeric(Gt/SsT),
              Pstl=as.numeric(stl/SsL),
              Pstrr=as.numeric(strr/Ssr),
              Pstrt=as.numeric(strt/SsT),
              Pstro=as.numeric(stro/SsO),
              PdlSl=as.numeric(dlSl/flux300yr$Cl),
              PdrSr=as.numeric(drSr/flux300yr$Cr),
              PdsSo=as.numeric(dsSo/flux300yr$Cs),
              PSr=as.numeric(senrTgdw/flux300yr$Br),
              PSl=as.numeric(senlTgdw/flux300yr$Bl),
              PSo=as.numeric(senoTgdw/Bo),
              PSt=0.001,
              PdwSo=as.numeric(dwSo/flux300yr$Cs)
)

ratesPtadea=as.data.frame(t(ratesPtadea)) ### these are the parameters to run the model in Fig 2 and obtain the results for P. taeda shown in the manuscript. These parameters were used in the phyton script (P.taeda_ages_and_transit_times.ipynb) for the computation of ages and transit times. 
```

## Sensitivity Analysis 

For the sensitivity analysis we used the parameters calculated above.

```{r}

ratesPtadea2=c(Rm=as.numeric(flux300yr$Rm/E),
              Fl=as.numeric(PFlT/E),
              Rl=as.numeric(PFrT/E),
              Sl=as.numeric(PFtT/E),
              BRl=as.numeric(PFoT/E),
              Gf=as.numeric(Gl/SsL),
              Gr=as.numeric(Gr/Ssr),
              Gbr=as.numeric(Go/SsO),
              Gs=as.numeric(Gt/SsT),
              Sf=as.numeric(stl/SsL),
              Sr=as.numeric(strr/Ssr),
              Ss=as.numeric(strt/SsT),
              Sbr=as.numeric(stro/SsO),
              Cf=as.numeric(dlSl/flux300yr$Cl),
              Cr=as.numeric(drSr/flux300yr$Cr),
              Cs=as.numeric(dsSo/flux300yr$Cs),
              Lr=as.numeric(senrTgdw/flux300yr$Br),
              Lf=as.numeric(senlTgdw/flux300yr$Bl),
              Lbr=as.numeric(senoTgdw/Bo),
              Ls=0.001,
              LSs=as.numeric(dwSo/flux300yr$Cs)
)
```

Load the functions for sensitivity and uncertainty analysis 
```{r}
source("/Users/_dherrera/NSC_ages_and_transit_times/NSC_ages_and_transit_times/code/script_function_senes_uncert_ACGCA.R")
```

The sensititvity analysis based on the Elementary Effect (Morrist 1991) was run using the function morris from the FME package  
```{r}

sensmeanage1 <- morris(model = systageogleV2, factors = names(ratesPtadea2), r = 100, binf = ratesPtadea2*0.1,
                       bsup =ratesPtadea2*2.5, 
                       design = list(type = "oat", levels = 150, grid.jump = 50))

mu <- apply(sensmeanage1$ee, 3, function(M){
  apply(M, 2, mean)
})

mu.star <- apply(abs(sensmeanage1$ee), 3, function(M){
  apply(M, 2, mean)
})
sigma <- apply(sensmeanage1$ee, 3, function(M){
  apply(M, 2, sd)
})

```

Plots for the sensitivity analysis
```{r}
#sensitivity for the mean age

png("Ptaeda_Mean_Age.png", width=400, height =400)
par(mar=c(5, 5,4,4))
plot(mu.star[,1], sigma[,1], main= "Mean Age",
     xlab=expression(paste(mu^"*")), ylab = expression(sigma),
     ylim=c(0, max(sigma[,1])+2),xlim=c(0, max(mu.star[,1])+2),
     cex=2, cex.axis=2, cex.lab=2, cex.main=3, pch=18 )
text(mu.star[,1], sigma[,1], row.names(mu),pos=3, cex=1.3)
dev.off()


png("Ptaeda_Mean_transit_time.png", width=400, height =400)
par(mar=c(5, 5,4,4))
plot(mu.star[,2], sigma[,2], main= "Mean Transit Time",
     xlab=expression(paste(mu^"*")), ylab = expression(sigma),
     ylim=c(0, max(sigma[,2]+2)),xlim=c(0, max(mu.star[,2])+0.1),
     cex=2, cex.axis=2, cex.lab=2, cex.main=3, pch=18 )
text(mu.star[,2], sigma[,2], row.names(mu),pos=3, cex=1.3)
dev.off()


```

plots for the trend of the sensitivity 
```{r}

### plots for the mean age
x=as.vector(sensmeanage1$X[,"LSs"])
y=sensmeanage1$y[,1]

m2=nls(y~x^z, start=list(z=0.1))
y2=sort(x)^-0.6005

par(mar=c(5,5,2,2))
png("Ptaeda_MeanAge_Vs_LSs", width = 350, height = 400)
par(mar=c(5,5,2,2))
plot(sensmeanage1$X[,"LSs"],sensmeanage1$y[,1], ylab="NSC system Man Age",
     xlab= "LSs", cex.axis=2, cex.lab=2, pch=18)
lines(sort(x), y2, col="red", lwd=2)
dev.off()

png("Ptaeda_MeanAge_Vs_Sbr", width = 350, height = 400)
par(mar=c(5,5,2,2))
plot(sensmeanage1$X[,"Sbr"], sensmeanage1$y[,1], ylab = "NSC system Man Age",
     xlab = "Sbr",  cex.axis=2, cex.lab=2, pch=18)
abline(lm(sensmeanage1$y[,1] ~ sensmeanage1$X[,"Sbr"]),col="red", lwd=2)
dev.off()


### plots for mean transit time

xGbr=as.vector(sensmeanage1$X[,"Gbr"])
yGbr=sensmeanage1$y[,2]

mGbr=nls(yGbr~b*xGbr^a, start=list(b=1,a=1))
yGbr2=2.3499*sort(xGbr)^-0.6271

png("Ptaeda_Meantransittime_Vs_Gbr", width = 350, height = 400)
par(mar=c(5,5,2,2))
plot(sensmeanage1$X[,"Gbr"], sensmeanage1$y[,2], ylab = "NSC system Mean Transit Time",
     xlab = "Gbr",  cex.axis=2, cex.lab=2, pch=18)
lines(sort(xGbr), yGbr2, col="red", lwd=2)
dev.off()

png("Ptaeda_Meantransittime_Vs_Sbr", width = 350, height = 400)
par(mar=c(5,5,2,2))
plot(sensmeanage1$X[,"Sbr"], sensmeanage1$y[,2], ylab = "NSC system Mean Transit Time",
     xlab = "Sbr",  cex.axis=2, cex.lab=2, pch=18)
abline(lm(sensmeanage1$y[,2] ~ sensmeanage1$X[,"Sbr"]),col="red", lwd=2)
dev.off()


```


## Uncertainty analysis 


Here we defined the range of variability of the parameters for the ACGCA model for P. taeda. This range was extracted from the supplementary inforamtion provided by Ogle and Pacala 2009. 

```{r}
paramogle_range=data.frame(Crwparam=c(0.308, 0.300, 0.316),
                     Radeff=c(4.5, 4, 5),
                     Pmax=c(2100, 2100, 2100),
                     Cgl=c(1.51, 1.44, 1.59),
                     Cgr=c(1.30, 1.22, 1.38),
                     Crw=c(1.47, 1.40, 1.54),
                     Rml=c(0.95, 0.47, 1.43),
                     Rmr=c(0.75, 0.33, 1.17),
                     Rms=c(0.025,0.017, 0.033),
                     Sl=c(0.33, 0.2, 0.5),
                     So=c(0.05, 0.02, 0.09),
                     Sr=c(0.5, 0.25, 1),
                     dl=c(0.11, 0.06, 0.16),
                     dr=c(0.08, 0.03, 0.13),
                     Stmaxspw=c(265000, 200000, 330000),
                     denspw=c(6.67e-7, 6.67e-7, 6.67e-7),
                     Xylspwrate=c(0.62, 0.59, 0.65),
                     k=c(0.55, 0.36, 0.64),
                     Maxcrwdepth=c(0.95, 0.95, 0.95),
                     pa=c(380000, 350000, 410000),
                     pb=c(380000, 350000, 410000),
                     pr=c(200000, 120000, 280000),
                     rootrad=c(0.00027, 0.00018, 0.00036),
                     SLA=c(0.0060, 0.0052, 0.0069),
                     n=c(0.71, 0.62, 0.8),
                     nb=c(0.045, 0.016, 0.066),
                     fiH=c(220, 196, 225),
                     lamb=c(0.95, 0.90, 1),
                     Hmax=c(42, 40, 44),
                     SWmax=c(0.06, 0.04, 0.08),
                     f1=c(4, 2, 6),
                     f2=c(2100, 1700, 2500), 
                     lambs=c(0.95, 0.90, 1),
                     lambh=c(0.95, 0.90, 1),
                     R0=c(1.434, 1.299, 1.669),
                     R40=c(3.873, 3.013,  4.730),
                     m=c(0.95, 0.95, 0.95))

rownames(paramogle_range)=c("mean", "min", "max")

```


We prepared the parameters to run the function modCRL for the Monte Carlo Simulations (MCS) to propagate the uncertainty from the ACGCA model parameters provided by Ogle and Pacala 2009 to the parameter of the compartmental model presented in figure 2 of te manuscript. This simulation can take hours, so we provide the resuls of our simulation in the file below "MSC_fluxes_Ptaed_steadystate3.Rda"

```{r}
pars1_l=list()
for(i in 1:ncol(paramogle_range)){
  pars1_l[i]=paramogle_range[1,i]
}
names(pars1_l)=names(paramogle_range)

parRanges=t(paramogle_range[c(2,3),])

# we choose the independent parameters to propagate the uncertainty 
# MCS_for_fluxes_steady_sate_Arub = modCRL(func = fluxes_steadystate_ogly,
#                                       parms = pars1_l,
#                                       parRange = parRanges[c(4:14,17,23:24,30),])
# 
##The results of the function can be loaded as 
load("MCS_fluxes_Ptaeda_steadystate3.Rda")

```

With this MCS we calculated the parameter mean, variance and standard deviation for the parameters of the model presented in the fig 2 of our manuscript, and then run another MCS with this parameter values to propagate the uncetainty to our model output: Mean ages and mean transit time

```{r}

variance_pars=apply(MCS_for_fluxes_steady_sate3, 2, var)
mean_pars=apply(MCS_for_fluxes_steady_sate3, 2, mean) 
sd_pars=apply(MCS_for_fluxes_steady_sate3, 2, sd)
min_pars=apply(MCS_for_fluxes_steady_sate3, 2, min)

flux_steady_state=as.data.frame(t(mean_pars))

### we calculated the transfer coefficients of the model in figure 2 based in the mean values of the parametes given by the ACGCA model. Same parameters given in the manuscript. 
transfer_coef_mean_Ptaeda=transfer_coeficients_calculation(flux_steady_state)

### we calcualted the parameters set of the model in Fig.2 corresponding to each realization of the MCS ran for the ACGCA model.  
transfer_coef_simulations=data.frame(matrix(nrow = 1, ncol = 21))
names_manuscript=c("Rm",  "Fl",  "Rl",  "Sl",  "BRl", "Gf",  "Gr",  "Gbr", "Gs",  "Sf",
                   "Sr",  "Ss",  "Sbr", "Cf",  "Cr", "Cs",  "Lr",  "Lf",  "Lbr", "Ls",  "LSs")
for (i in 1:nrow(MCS_for_fluxes_steady_sate3)){
  transfer_coef_simulations[i,]=transfer_coeficients_calculation(MCS_for_fluxes_steady_sate3[i,])
}
names(transfer_coef_simulations)=names_manuscript


## we estimated then the mean, the variance and the sd for the transfer coeficients of the model in fig 2

mean_transfer_coeficients=apply(transfer_coef_simulations,2,mean)
var_tansfer_coeficients=apply(transfer_coef_simulations, 2, var)
sd_transfer_coeficients=apply(transfer_coef_simulations, 2, sd)

```


Now we progapate this uncertainties to our model output. This simulation can take hours, so we provide the results of the simulation in the file "MCS_meanage_Ptaeda.rda"

```{r}
## we prepare again the list of model parameters to run the modCRK function 

pars1_l2=list()
for(i in 1:length(mean_transfer_coeficients)){
  pars1_l2[i]=mean_transfer_coeficients[i]
}
names(pars1_l2)=names(mean_transfer_coeficients)

## we ran the MCS for the parameters that had the most influential effect in the model output, which were identified in the sensitivity analysis. 

# MCS_meanage_Arubrum <- modCRL(func = systageogleV3, parms = pars1_l2, parMean = mean_transfer_coeficients[c(21,16,8,13)],
#                              parCovar =diag(var_tansfer_coeficients[c(21,16,8,13)]),
#                              dist = "norm", num = 1000)

## the results for this MCS can be loaded
load("MCS_meanage_Ptaeda.rda")

## the standard statistics can be calcualted as
sd_Ptaeda_meanage=apply(MCS_meanage_Ptaeda,2,sd)
names(sd_Ptaeda_meanage)=names(MCS_meanage_Ptaeda)
mean_Ptaeda_meanage=apply(MCS_meanage_Ptaeda,2,mean)

```

With these resuls we built the plot that is part of the supplementary information of the manuscript (Fig 1sup). 

```{r}
png("Ptaeda_meanSysAgebypool_uncertianty.png",
        width = 900,
        height = 700)
par(mfrow=c(2,5), mar=c(5,5,2,2))
hist(MCS_meanage_Ptaeda$meansystemage, main = "", xlab = "Mean System Age", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$meantransittime, main = "", xlab = "Mean Transit Time", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$E, main = "", xlab = "E", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$FLNSC, main = "", xlab = "FLNSC", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$RLNSC, main = "", xlab = "RLNSC", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$BRLNSC, main = "", xlab = "BRLNSC", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$SLNSC, main = "", xlab = "SLNSC", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$FSNSC, main = "", xlab = "FSNSC", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$RSNSC, main = "", xlab = "RSNSC", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Ptaeda$SSNSC, main = "", xlab = "SSNSC", breaks = 50, cex.axis=2, cex.lab=2)
dev.off()

```
