---
title: "A.rubrum_snesitivity_and_uncertainty_analysis"
author: "David_Herrera"
date: "10/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document present the calculations related to the parameter values estimation, uncertainty and sensitivity analysis for the model corresponting to A. rubrum, detailes in the Figure 2 of the manuscript "Probability distributions of non-structural carbon ages and transit times provide insights on carbon allocation dynamics of mature trees" 

### packages needed 
```{r}
library(SoilR)
library(boot)
library(FME) 
library(sensitivity)
```

set the directory where all the auxiliary files are stored. The functions for running the ACGCA model are in the sicript "script_functions_model_ogleV2.R", This script should be sourced.  

```{r}
setwd("/Users/_dherrera/NSC_ages_and_transit_times/NSC_ages_and_transit_times/code/A_rubrum/")
source("/Users/_dherrera/NSC_ages_and_transit_times/NSC_ages_and_transit_times/code/script_functions_model_ACGCA.R")

```

## A. rubrum parameter values computations 

We wrote down the parameters provided for Ogle and Pacala (2009) and placed them in a data frame called paramogle. These parameters are used for running the model ACGCA and estimating the parameters for the model presented in the mentioned figure 2 of the manuscript. the data frame was contructed with the mean values for P. taeda (first row) and A. rubrum (second row). From here on we will just use the parameters corresponding to A. rubrum. 



```{r}

paramogle=data.frame(Crwparam=c(0.308, 0.365),
                     Radeff=c(4.5, 6.75),
                     Pmax=c(2100, 2100),
                     Cgl=c(1.51, 1.45),
                     Cgr=c(1.30, 1.25),
                     Crw=c(1.47, 1.37),
                     Rml=c(0.95, 1.25),
                     Rmr=c(0.75, 0.75),
                     Rms=c(0.025, 0.025),
                     Sl=c(0.33, 1),
                     So=c(0.05, 0.05),
                     Sr=c(0.5, 1),
                     dl=c(0.11, 0.095),
                     dr=c(0.08, 0.15),
                     Stmaxspw=c(265000, 131000),
                     denspw=c(6.67e-7, 6.67e-7),
                     Xylspwrate=c(0.62, 0.12),
                     k=c(0.55, 0.7),
                     Maxcrwdepth=c(0.95, 0.95),
                     pa=c(380000, 525000),
                     pb=c(380000, 525000),
                     pr=c(200000, 160000),
                     rootrad=c(0.00027, 0.00015),
                     SLA=c(0.0060, 0.0141),
                     n=c(0.71, 0.64),
                     nb=c(0.045, 0.045),
                     fiH=c(220, 263),
                     lamb=c(0.95, 0.95),
                     Hmax=c(42, 27.5),
                     SWmax=c(0.06, 0.10),
                     f1=c(4, 4),
                     f2=c(2100, 7000), 
                     lambs=c(0.95, 0.95),
                     lambh=c(0.95, 0.95),
                     R0=c(1.434, 1.909),
                     R40=c(3.873, 5.592),
                     m=c(0.95, 0.95))

paraPtadea=paramogle[2,] ### here we choosed the parameters for A. rubrum. 
```


Then we ran the ACGCA model for 800 years (niter) to assure that we selected the parameters corresponding to a steady state condition for A. rubrum. This simulation can take time, therefore we provide the result in the loaded file "respArubrum_Pmax.Rda"

```{r}


 # r1=0.0001### inital value for radious
 # respPtadea=ACGCAtargetallom(r1=r1, dr=0, t=0, paraPtadea=paraPtadea, Cs=c(0))## initial conditions 
 # respPtadea4=modhealty(initcond = respPtadea, niter=800, param=paraPtadea, tao=0.000001, errori=10000)
 # respPtadea4=respPtadea4[-c(nrow(respPtadea4)),] ## the last row is incomplete then we eliminated.


 load("respArubrum_PmaxV6.Rda")# The results of the previous simulation can be laded here. 
 
```

Then we calculated the fluxes, the pool sizes and the stransfer coeficients for the model in Figure 2

```{r}

## We took then the fluxes and stocks in stady state: the values at the point of 700 years of the simulations of the ACGCA model

flux300yr=respPtadea4[which(respPtadea4$t==700),] 


## With these values we proceeded to calculate the transfer coeficients for the model used in the manuscript and described in Figure 2

E=flux300yr$E+flux300yr$Rm ##g of glucose 

## fluxes for leave compartment 

PFlT=flux300yr$Fl*flux300yr$E; ## PFlT is the amount of g glucose allocated to leaves 
SsL=PFlT # SsL is the fast cycling pool of sugars in the leaves
Gl=PFlT*(paraPtadea["Cgl"]/(paraPtadea["Cgl"]+paraPtadea["dl"])); Gl### grams of glucose per year invested in constructing leaves (flux of glucose to grow leaves)
FB=flux300yr$Bl## biomass (g dw) in the leaves 
biomlgdw=PFlT/(paraPtadea["Cgl"]+paraPtadea["dl"]); biomlgdw## grams of wood constructed from the allocated carbon to leaves
stl=PFlT*(paraPtadea["dl"]/(paraPtadea["Cgl"]+paraPtadea["dl"])); stl## grams of glucose per year stored in the leaves 
senlTgdw=paraPtadea["Sl"]*flux300yr$Bl; senlTgdw## g of wood lost in litter fall per year 
senlT=paraPtadea["Sl"]*flux300yr$Bl*paraPtadea["Cgl"] ## g of glucose per year lost in the litter fall 
dlSl=paraPtadea["Sl"]*flux300yr$Bl*paraPtadea["dl"]; dlSl ##g of glucose per year translocated to the E pool from the storage
FSNSC=flux300yr$Cl ## size of the storage pool in the leaves 


## fluxes for fine root compartment 
PFrT=flux300yr$Fr*flux300yr$E; PFrT## amount of g of glucose allocated to roots per year 
Ssr=PFrT ## g glucose in the fast cycling pool of sugar in the roots
Gr=PFrT*(paraPtadea["Cgr"]/(paraPtadea["Cgr"]+paraPtadea["dr"])); Gr## grams of glucoes allocated to biomass in roots
biomrgdw=PFrT/(paraPtadea["Cgr"]+paraPtadea["dr"]); biomrgdw## grams of dry wood constructed from the allocated carbon to roots
strr=PFrT*(paraPtadea["dr"]/(paraPtadea["Cgr"]+paraPtadea["dr"])); strr## grams of glucose per year stored in the roots 
senrTgdw=paraPtadea["Sr"]*flux300yr$Br## g of wood lost in senescence of fine roots 
senrT=paraPtadea["Sr"]*flux300yr$Br*paraPtadea["Cgr"] ## g of glucose lost in senescence of fine roots 
drSr=paraPtadea["Sr"]*flux300yr$Br*paraPtadea["dr"];drSr ## g of glucose reallocate to the transient comparment (E) from roots 
RSNSC=flux300yr$Cr## size of the sotrage compartment in roots


## fluxes for woody compartments 

PFtT=flux300yr$Ft*flux300yr$E; PFtT## amount of g glucose allocated to the main trunk
SsT=PFtT## g of glucose allocated to the fast cycling pool in the trunk 
Gt=PFtT*(paraPtadea["Crw"]/(paraPtadea["Crw"]+flux300yr$ds))#g of glucose per year allocated to grow wood in the trunk 
SB=flux300yr$Bts+flux300yr$Bth ## g of wood in the trunk (biomase of the main trunk)
strt=PFtT*(flux300yr$ds/(paraPtadea["Crw"]+flux300yr$ds))## grams of glucose per year allocated to storage in the trunk 

## fluxes for other wood compartment (branches and coarse roots)
PFoT=flux300yr$Fo*flux300yr$E; PFoT## amount of g glucose allocated to branches and coarse roots per year 
SsO=PFoT ## size of the fast cycling pool of branches and coarse roots  in g of glucose
Go=PFoT*(paraPtadea["Crw"]/(paraPtadea["Crw"]+flux300yr$ds))### grams of glucose per year invested in constructing branches and coarse roots
Bo=flux300yr$Bos+flux300yr$Boh ## g of wood in the branches and coarse root
senoTgdw=PFoT/(paraPtadea["Crw"]+flux300yr$ds)#g of wood per year lost in litter fall
senoT=PFoT/(paraPtadea["Crw"]+flux300yr$ds)*paraPtadea["Crw"]## g of glucose lost in litter fall
stro=PFoT*(flux300yr$ds/(paraPtadea["Crw"]+flux300yr$ds))## g of glucose per year allocated to storage inthe wood
So=(PFoT/(paraPtadea["Crw"]+flux300yr$ds))/Bo## g of wood per year lost in litter fall from branches and coarse roots 
dsSo=flux300yr$ds*So*flux300yr$Bos### g of glucose per year retranslocated to E from branches and coarse roots 
dwSo=strt+stro-dsSo## g of gluc per year from the carbon sotorage compartment lost in litterfall  
SSNSC=flux300yr$Cs## glocose stored in wood of branches and coarse roots

### calculation of the carbon transfer coefficients, the parameters of our model described in Fig.2.  

ratesPtadea=c(Rm=as.numeric(flux300yr$Rm/E),
              Fl=as.numeric(PFlT/E),
              Fr=as.numeric(PFrT/E),
              Ft=as.numeric(PFtT/E),
              Fo=as.numeric(PFoT/E),
              PGl=as.numeric(Gl/SsL),
              PGr=as.numeric(Gr/Ssr),
              PGo=as.numeric(Go/SsO),
              PGt=as.numeric(Gt/SsT),
              Pstl=as.numeric(stl/SsL),
              Pstrr=as.numeric(strr/Ssr),
              Pstrt=as.numeric(strt/SsT),
              Pstro=as.numeric(stro/SsO),
              PdlSl=as.numeric(dlSl/flux300yr$Cl),
              PdrSr=as.numeric(drSr/flux300yr$Cr),
              PdsSo=as.numeric(dsSo/flux300yr$Cs),
              PSr=as.numeric(senrTgdw/flux300yr$Br),
              PSl=as.numeric(senlTgdw/flux300yr$Bl),
              PSo=as.numeric(senoTgdw/Bo),
              PSt=0.001,
              PdwSo=as.numeric(dwSo/flux300yr$Cs)
)

ratesPtadea=as.data.frame(t(ratesPtadea)) ### this are the parameters to run the model in Fig 2 and obtain the results for A. rubrum shown in the manuscript. These parameters were used in the phyton script (A.rubrum_ages_and_transit_times.ipynb) for the computation of ages and transit times. 
```

## Sensitivity Analysis 

For the sensitivity analysis we used the parameters calculated above and we specified a wide variability range to see the behavior of the output model in a wide range of parameters.

```{r}
ratesPtadea2=c(Rm=as.numeric(flux300yr$Rm/E),
              Fl=as.numeric(PFlT/E),
              Rl=as.numeric(PFrT/E),
              Sl=as.numeric(PFtT/E),
              BRl=as.numeric(PFoT/E),
              Gf=as.numeric(Gl/SsL),
              Gr=as.numeric(Gr/Ssr),
              Gbr=as.numeric(Go/SsO),
              Gs=as.numeric(Gt/SsT),
              Sf=as.numeric(stl/SsL),
              Sr=as.numeric(strr/Ssr),
              Ss=as.numeric(strt/SsT),
              Sbr=as.numeric(stro/SsO),
              Cf=as.numeric(dlSl/flux300yr$Cl),
              Cr=as.numeric(drSr/flux300yr$Cr),
              Cs=as.numeric(dsSo/flux300yr$Cs),
              Lr=as.numeric(senrTgdw/flux300yr$Br),
              Lf=as.numeric(senlTgdw/flux300yr$Bl),
              Lbr=as.numeric(senoTgdw/Bo),
              Ls=0.001,
              LSs=as.numeric(dwSo/flux300yr$Cs)
)

```

Load the functions for sensitivity and uncertainty analyses 
```{r}
source("/Users/_dherrera/NSC_ages_and_transit_times/NSC_ages_and_transit_times/code/script_function_senes_uncert_ACGCA.R")
```

The sensitivity analysis based on the Elementary Effect (Morrist 1991) was run using the function morris from the FME package  
```{r}

sensmeanage1 <- morris(model = systageogleV2, factors = names(ratesPtadea2), r = 100, binf = ratesPtadea2*0.1,
                       bsup =ratesPtadea2*2.5, 
                       design = list(type = "oat", levels = 150, grid.jump = 50))

mu <- apply(sensmeanage1$ee, 3, function(M){
  apply(M, 2, mean)
})

mu.star <- apply(abs(sensmeanage1$ee), 3, function(M){
  apply(M, 2, mean)
})
sigma <- apply(sensmeanage1$ee, 3, function(M){
  apply(M, 2, sd)
})

```

Plots for the sensitivity analysis
```{r}
#sensitivity for the mean age

png("Arubrum_Mean_Age.png", width=400, height =400)
par(mar=c(5, 5,4,4))
plot(mu.star[,1], sigma[,1], main= "Mean Age",
     xlab=expression(paste(mu^"*")), ylab = expression(sigma),
     ylim=c(0, max(sigma[,1])+2),xlim=c(0, max(mu.star[,1])+2),
     cex=2, cex.axis=2, cex.lab=2, cex.main=3, pch=18 )
text(mu.star[,1], sigma[,1], row.names(mu),pos=3, cex=1.3)
dev.off()

png("Arubrum_Mean_transit_time.png", width=400, height =400)
par(mar=c(5, 5,4,4))
plot(mu.star[,2], sigma[,2], main= "Mean Transit Time",
     xlab=expression(paste(mu^"*")), ylab = expression(sigma),
     ylim=c(0, max(sigma[,2]+2)),xlim=c(0, max(mu.star[,2])+0.1),
     cex=2, cex.axis=2, cex.lab=2, cex.main=3, pch=18 )
text(mu.star[,2], sigma[,2], row.names(mu),pos=3, cex=1.3)
dev.off()

```


plots for the trend of the sensitivity 
```{r}

### plots for the mean age
x=as.vector(sensmeanage1$X[,"LSs"])
y=sensmeanage1$y[,1]

m2=nls(y~x^-z, start=list( z=1))
y2=sort(x)^-0.5966

par(mar=c(5,5,2,2))
png("Arubrum_MeanAge_Vs_LSs", width = 350, height = 400)
par(mar=c(5,5,2,2))
plot(sensmeanage1$X[,"LSs"],sensmeanage1$y[,1], ylab="NSC system Man Age",
     xlab= "LSs", cex.axis=2, cex.lab=2, pch=18)
lines(sort(x), y2, col="red", lwd=2)
dev.off()

png("Arubrum_MeanAge_Vs_Sbr", width = 350, height = 400)
plot(sensmeanage1$X[,"Sbr"], sensmeanage1$y[,1], ylab = "NSC system Man Age",
     xlab = "Sbr",  cex.axis=2, cex.lab=2, pch=18)
abline(lm(sensmeanage1$y[,1] ~ sensmeanage1$X[,"Sbr"]),col="red", lwd=2)
dev.off()


### plots for mean transit time

xGbr=as.vector(sensmeanage1$X[,"Gbr"])
yGbr=sensmeanage1$y[,2]

mGbr=nls(yGbr~b*xGbr^a, start=list(b=1,a=1))
yGbr2=2.92*sort(xGbr)^-0.6098

png("Arubrum_Meantransittime_Vs_Gbr", width = 350, height = 400)
par(mar=c(5,5,2,2))
plot(sensmeanage1$X[,"Gbr"], sensmeanage1$y[,2], ylab = "NSC system Mean Transit Time",
     xlab = "Gbr",  cex.axis=2, cex.lab=2, pch=18)
lines(sort(xGbr), yGbr2, col="red", lwd=2)
dev.off()

png("Arubrum_Meantransittime_Vs_Sbr", width = 350, height = 400)
par(mar=c(5,5,2,2))
plot(sensmeanage1$X[,"Sbr"], sensmeanage1$y[,2], ylab = "NSC system Mean Transit Time",
     xlab = "Sbr",  cex.axis=2, cex.lab=2, pch=18)
abline(lm(sensmeanage1$y[,2] ~ sensmeanage1$X[,"Sbr"]),col="red", lwd=2)
dev.off()


```


## Uncertainty analysis 

We defined the range of variability of the parameters for the ACGCA model for A. rubrum. This ranged were extracted from the supplementary inforamtion provided by Ogle and Pacala 2009. 

```{r}
paramogle_range=data.frame(Crwparam=c(0.365, 0.360, 0.370),
                           Radeff=c(6.75, 6, 7.5),
                           Pmax=c(2100, 2100, 2100),
                           Cgl=c(1.45, 1.41, 1.49),
                           Cgr=c(1.25, 1.22, 1.27),
                           Crw=c(1.37, 1.35, 1.39),
                           Rml=c(1.25, 0.83, 1.66),
                           Rmr=c(0.75, 0.27, 1.23),
                           Rms=c(0.025, 0.012, 0.038),
                           Sl=c(1, 1, 1),
                           So=c(0.05, 0.02, 0.09),
                           Sr=c(1, 0.9, 1.1),
                           dl=c(0.095, 0.086, 0.104),
                           dr=c(0.15, 0.079, 0.22),
                           Stmaxspw=c(131000, 100000, 162000),
                           denspw=c(6.67e-7, 6.67e-7, 6.67e-7),
                           Xylspwrate=c(0.12, 0.11, 0.13),
                           k=c(0.7, 0.6, 0.8),
                           Maxcrwdepth=c(0.95, 0.95, 0.95),
                           pa=c(525000, 460000, 585000),
                           pb=c(525000, 460000, 585000),
                           pr=c(160000, 60000 ,260000),
                           rootrad=c(0.00015, 0.00010, 0.00020),
                           SLA=c(0.0141, 0.0119, 0.0165),
                           n=c(0.64, 0.64, 0.64),
                           nb=c(0.045,0.045, 0.045),
                           fiH=c(263, 234, 293),
                           lamb=c(0.95, 0.90, 1),
                           Hmax=c(27.5, 25.5, 29.5),
                           SWmax=c(0.10, 0.05, 0.15),
                           f1=c(4, 2, 6),
                           f2=c(7000, 5000, 9000), 
                           lambs=c(0.95, 0.90, 1),
                           lambh=c(0.95, 0.90, 1),
                           R0=c(1.909, 1.674, 2.144),
                           R40=c(5.592, 4.732, 6.451),
                           m=c(0.95, 0.95, 0.95))

rownames(paramogle_range)=c("mean", "min", "max")

```


We prepared the parameters to run the function modCRL for the Monte Carlo Simulations (MCS) and run the model to propagate the uncertainty from the ACGCA model parameters provided by Ogle and Pacala 2009 to the parameters of the compartmental model presented in the figure 2 of te manuscript. 

```{r}
pars1_l=list()
for(i in 1:ncol(paramogle_range)){
  pars1_l[i]=paramogle_range[1,i]
}
names(pars1_l)=names(paramogle_range)

parRanges=t(paramogle_range[c(2,3),])

# MCS_for_fluxes_steady_sate_Arub = modCRL(func = fluxes_steadystate_ogly,
#                                       parms = pars1_l,
#                                       parRange = parRanges[c(4:14,17,23:24,30),])
# 
##The results of the function can be loaded as 
load("MCS_fluxes_steadystate_Arub.Rda")

## we renamed it
MCS_for_fluxes_steady_sate3=MCS_for_fluxes_steady_sate_Arub
```

With this MCS we calculated the parameter mean, variance and standard deviation for the model presented in the fig 2 of our manuscript, and then run another MCS with this parameter values to propagate the uncetainty to our model output: Mean ages and mean transit time


```{r}

variance_pars=apply(MCS_for_fluxes_steady_sate3, 2, var)
mean_pars=apply(MCS_for_fluxes_steady_sate3, 2, mean) 
sd_pars=apply(MCS_for_fluxes_steady_sate3, 2, sd)
min_pars=apply(MCS_for_fluxes_steady_sate3, 2, min)

flux_steady_state=as.data.frame(t(mean_pars))

### we calculated the transfer coeficients of the model in figure 2 based in the mean values of the parametes given by the ACGCA model. Same parameters given in the manuscript. 
transfer_coef_mean_Arubrum=transfer_coeficients_calculation(flux_steady_state)

###we calcualted the parameters set of the model in Fig.2 corresponding to each realization of the MCS ran for the ACGCA model. 
transfer_coef_simulations=data.frame(matrix(nrow = 1, ncol = 21))
names_manuscript=c("Rm",  "Fl",  "Rl",  "Sl",  "BRl", "Gf",  "Gr",  "Gbr", "Gs",  "Sf",
                   "Sr",  "Ss",  "Sbr", "Cf",  "Cr", "Cs",  "Lr",  "Lf",  "Lbr", "Ls",  "LSs")
for (i in 1:nrow(MCS_for_fluxes_steady_sate3)){
  transfer_coef_simulations[i,]=transfer_coeficients_calculation(MCS_for_fluxes_steady_sate3[i,])
}
names(transfer_coef_simulations)=names_manuscript


## we estimated then the mean, the variance and the sd for the transfer coeficients of the model in fig 2

mean_transfer_coeficients=apply(transfer_coef_simulations,2,mean)
var_tansfer_coeficients=apply(transfer_coef_simulations, 2, var)
sd_transfer_coeficients=apply(transfer_coef_simulations, 2, sd)
```


Now we progapate this uncertainties to our model output

```{r}

## we prepared again the list of model parameters to run the modCRK function 

pars1_l2=list()
for(i in 1:length(mean_transfer_coeficients)){
  pars1_l2[i]=mean_transfer_coeficients[i]
}
names(pars1_l2)=names(mean_transfer_coeficients)

## we ran the MCS for the parameters that had the most influential effect in the model output, which were identified in the sensitivity analysis. 

# MCS_meanage_Arubrum <- modCRL(func = systageogleV3, parms = pars1_l2, parMean = mean_transfer_coeficients[c(21,16,8,13)],
#                              parCovar =diag(var_tansfer_coeficients[c(21,16,8,13)]),
#                              dist = "norm", num = 1000)

## the results for this MCS can be loaded
load("MCS_meanages_Arubrim.rda")

## the standard statistics can be calcualted as
sd_Arub_meanage=apply(MCS_meanage_Arubrum,2,sd)
names(sd_Arub_meanage)=names(MCS_meanage_Arubrum)
mean_Arub_meanage=apply(MCS_meanage_Arubrum,2,mean)

```

With these resuls we built the plot that is part of the supplementary information of the manuscript. 

```{r}
png("Arubrum_meanSysAgebypool_uncertianty.png",
    width = 900,
    height = 700)
par(mfrow=c(2,5), mar=c(5,5,2,2))
hist(MCS_meanage_Arubrum$meansystemage, main = "", xlab = "Mean System Age", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$meantransittime, main = "", xlab = "Mean Transit Time", breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$E, main = "", xlab = "E",breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$FLNSC, main = "", xlab = "FLNSC",breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$RLNSC, main = "", xlab = "RLNSC",breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$BRLNSC, main = "", xlab = "BRLNSC",breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$SLNSC, main = "", xlab = "SLNSC",breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$FSNSC, main = "", xlab = "FSNSC",breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$RSNSC, main = "", xlab = "RSNSC",breaks = 50, cex.axis=2, cex.lab=2)
hist(MCS_meanage_Arubrum$SSNSC, main = "", xlab = "SSNSC",breaks = 50, cex.axis=2, cex.lab=2)
dev.off()

```
